---
title: "Blast Adult ERP Analysis"
author: "JM Schneider"
date: "04/16/2021"
output: html_document
---

## This code analyzes the BLAST projects' adult EEG data.

### Import libraries
```{r, echo=FALSE, include=FALSE}
library(psych);
library(readr);
library(corrplot);
library(reshape);
library(lmerTest);
library(ggplot2);
library(scales);
library(ggbeeswarm);
library(Hmisc);
library(dplyr);
```
### Import data as dataframe
```{r}
df_corr=read.csv('blast_online_erp_adult_corr_data.csv')
df_rt_slope=read.csv('blast_online_adult_ssl_indiv_rts.csv')
df_factor=read.csv('blast_adult_factor_analysis_n45.csv')

df_corr$mean_rt <- NULL
df_corr$scaled_rt_slope <- NULL

df_factor$Local = as.factor(df_factor$Local)
df_factor$Global = as.factor(df_factor$Global)
df_factor$Domain = as.factor(df_factor$Domain)
```

```{r}
df_rt_slope <- df_rt_slope %>% 
  dplyr::rename(
    subject = par_id
  )
df_rt_slope = subset(df_rt_slope, select = c(subject,mean_rt,scaled_rt_slope) )
df_corr <- merge(df_corr,df_rt_slope, by = "subject")
```

## 3.2.2 Contribution of domain, global probability, and local probability
### LMER of MMN (Early ERP Data)
```{r, echo=TRUE}
df_factor$Local = relevel(df_factor$Local,"Short")
model<- lmer(Early~ Domain*Global*Local + (1+(Global*Local)|subject),data = df_factor,control = lmerControl(calc.derivs = FALSE))
summary(model)
```
### LMER of LDN/P3a (Late ERP Data)
```{r, echo=TRUE}
model<- lmer(Late~ Domain*Global*Local + (1+(Global*Local)|subject),data = df_factor,control = lmerControl(calc.derivs = FALSE))
summary(model)
```

### Posthoc pairwise t-tests used to gauge interaction effect 
```{r}
blast_high<-df_factor[which(df_factor$Global=="High"),]
blast_low<-df_factor[which(df_factor$Global=="Low"),]
blast_short<-df_factor[which(df_factor$Local=="Short"),]
blast_long<-df_factor[which(df_factor$Local=="Long"),]
pairwise.t.test(blast_short$Late, blast_short$Global, p.adj = 'bonferroni') #significant
pairwise.t.test(blast_long$Late, blast_long$Global, p.adj = 'bonferroni')
pairwise.t.test(blast_high$Late, blast_high$Local, p.adj = 'bonferroni') #significant
pairwise.t.test(blast_low$Late, blast_low$Local, p.adj = 'bonferroni')
```
## Plotting of LDN interaction effect
### Plot the line graphs to illustrate interaction between Local and Global Interaction in Late Time Window
```{r}
d <- df_factor %>%
       select(Late,Local,Global) %>%  # select relevant variables
       mutate(Local = factor(Local),
              Global = factor(Global))
head(d)
d %>% 
  group_by(Local, Global) %>%
  summarise(Late_mean = mean(Late))
sum_d <- d %>% 
          group_by(Local, Global) %>%
          summarise(Late_mean = mean(Late),
                    se   = sd(Late)/sqrt(n()))
sum_d
```

```{r}
pd <- position_dodge(width = 0)
sum_d %>%
  ggplot(aes(x = Global, y = Late_mean, group = Local)) +
    geom_line(aes(linetype = Local), position = pd) +
    geom_errorbar(aes(ymin = Late_mean - se, ymax = Late_mean + se),
                  width = .1, position = pd) +
    geom_point(size = 4, position = pd) +
    geom_point(size = 3, color = "white", position = pd) + 
    guides(linetype = guide_legend("Local")) +
    labs(x = "Global",
         y = "Average ERP Amplitude") +
  theme(
    text=element_text(size=20),
    panel.background = element_rect(fill = "white"),         # Set plot background to white
    legend.key  = element_rect(fill = "white"),              # Set legend item backgrounds to white
    axis.line.x = element_line(colour = "black", size = 1),  # Add line to x axis
    axis.line.y = element_line(colour = "black", size = 1)   # Add line to y axis
  )
```

## This section is used to analyze the correlation between web-based SSL and BLAST ERP amplitudes.

### Compare Correlation matrix and Factor matrix subjects
```{r, eval=TRUE}
df_factor_subset <- df_factor[1:45,]
df_factor_subset$subject[!(df_factor_subset$subject %in% df_corr$subject)]
df_corr$subject[!(df_corr$subject %in% df_factor_subset$subject)]
```

### Remove subjects that do not have SSL online data or do not have ERP data
```{r, eval=TRUE}
df_factor<- df_factor[!grepl("blast_a_012", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_020", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_027", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_029", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_031", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_036", df_factor$subject),]
df_factor<- df_factor[!grepl("blast_a_025", df_factor$subject),]
df_corr<- df_corr[!grepl("blast_a_017", df_corr$subject),]
df_corr<- df_corr[!grepl("blast_a_032", df_corr$subject),]
df_corr<- df_corr[!grepl("blast_a_035", df_corr$subject),]
df_corr<- df_corr[!grepl("blast_a_040", df_corr$subject),]
df_corr<- df_corr[!grepl("blast_a_052", df_corr$subject),]
df_corr<- df_corr[!grepl("blast_a_025", df_corr$subject),]
```

```{r, eval=TRUE}
df_corr$scaled_rt_slope = as.numeric(as.character(df_corr$scaled_rt_slope))

# Early Correlation values of interest
df_factor$Early = as.numeric(df_factor$Early)
df_early = cast(df_factor,subject~Domain+Global+Local,value="Early")

df_corr_early =df_corr

df_corr_early$Short = (df_early$Syllable_High_Short+df_early$Syllable_Low_Short+df_early$Voice_High_Short+df_early$Voice_Low_Short)/4
df_corr_early$Long = (df_early$Syllable_High_Long+df_early$Syllable_Low_Long+df_early$Voice_High_Long+df_early$Voice_Low_Long)/4

df_corr_early$Long_Short_MMN = df_corr_early$Long- df_corr_early$Short

# Late Correlation values of interest
df_factor$Late = as.numeric(df_factor$Late)
df_late = cast(df_factor,subject~Domain+Global+Local,value="Late")

df_corr_late =df_corr

df_corr_late$High = (df_late$Syllable_High_Short+df_late$Syllable_High_Long+df_late$Voice_High_Short+df_late$Voice_High_Long)/4
df_corr_late$Low = (df_late$Syllable_Low_Short+df_late$Syllable_Low_Long+df_late$Voice_Low_Short+df_late$Voice_Low_Long)/4
df_corr_late$Low_High_MMN = df_corr_late$Low-df_corr_late$High 

df_corr_late$Global_long =  ((df_late$Syllable_Low_Long+df_late$Voice_Low_Long)/2) - ((df_late$Syllable_High_Long+df_late$Voice_High_Long)/2)
df_corr_late$Global_short = ((df_late$Syllable_Low_Short+df_late$Voice_Low_Short)/2) - ((df_late$Syllable_High_Short+df_late$Voice_High_Short)/2)
df_corr_late$Local_high = ((df_late$Syllable_High_Long+df_late$Voice_High_Long)/2)- ((df_late$Syllable_High_Short+df_late$Voice_High_Short)/2)
df_corr_late$Local_low = ((df_late$Syllable_Low_Long+df_late$Voice_Low_Long)/2)-((df_late$Syllable_Low_Short+df_late$Voice_Low_Short)/2)


df_corr_early$Long_Short_MMN = as.numeric(df_corr_early$Long_Short_MMN)
df_corr_late$Global_long = as.numeric(df_corr_late$Global_long)
df_corr_late$Global_short = as.numeric(df_corr_late$Global_short)
df_corr_late$Local_high = as.numeric(df_corr_late$Local_high)
df_corr_late$Local_low = as.numeric(df_corr_late$Local_low)
df_corr_late$Low_High_MMN = as.numeric(df_corr_late$Low_High_MMN)
```

## 3.1 Statistical Learning Behavioral Results
### One Sample t-test of behavioral data
```{r}
t.test(df_corr$accuracy, mu = .50, alternative = "two.sided")
t.test(df_corr$scaled_rt_slope, mu = 0, alternative = "two.sided")
cor.test(df_corr$accuracy, df_corr$scaled_rt_slope,  method = "pearson") #marginally significant
```

## 3.3 Correlation Analysis
```{r}
# Early time window
cor.test(df_corr_early$Long_Short_MMN, df_corr_early$scaled_rt_slope,  method = "spearman", exact = FALSE) #significant

# Late time window
cor.test(df_corr_late$Low_High_MMN, df_corr_late$scaled_rt_slope,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Global_long, df_corr_late$scaled_rt_slope,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Global_short, df_corr_late$scaled_rt_slope,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Local_high, df_corr_late$scaled_rt_slope,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Local_low, df_corr_late$scaled_rt_slope,  method = "spearman", exact = FALSE) #significant

# Same effects and accuracy; early time window
cor.test(df_corr_early$Long_Short_MMN, df_corr_early$accuracy,  method = "spearman", exact = FALSE) 

# late time window
cor.test(df_corr_late$Low_High_MMN, df_corr_late$accuracy,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Global_long, df_corr_late$accuracy,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Global_short, df_corr_late$accuracy,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Local_high, df_corr_late$accuracy,  method = "spearman", exact = FALSE)
cor.test(df_corr_late$Local_low, df_corr_late$accuracy,  method = "spearman", exact = FALSE) 

cor.test(df_corr_late$Low_High_MMN, df_corr_early$Long_Short_MMN,  method = "pearson", exact = FALSE)
```

## Learner Groups: Relationship between the neural responses to prediction errors and individual learners’ statistical learning performance 

### Good and Poor Learners as determined by RT Slope
```{r}
good_learn<-df_corr[which(df_corr$scaled_rt_slope < 0),]
poor_learn<-df_corr[which(df_corr$scaled_rt_slope > 0),]

df_corr_late$learner2 = "P3"
df_corr_late[which(df_corr_late$Local_low < 0),]$learner2="LDN"
df_corr_late$learner2 = as.factor(df_corr_late$learner2)

t.test(scaled_rt_slope~learner2,data = df_corr_late, var.equal = TRUE)
t.test(accuracy~learner2,data = df_corr_late, var.equal = TRUE)
```

## Plot late correlation between P3 and RT Slope
```{r}
ggplot(df_corr_late, aes(x=Local_low, y=scaled_rt_slope)) + geom_point(size=3, color="navy blue", shape=19) + labs(x="P3 Amplitude (uV)\nLocal (Low vs. High), Low Frequency ", y = "Speech SL (RT Slope)") +  geom_smooth(method = "loess", se=FALSE) + theme_classic(base_size = 20.0) +
theme(plot.title = element_text(hjust = 0.5))
```

## Plot the demonstration of correlation between LDN/P3 in late time window and learning
```{r, echo=FALSE}
ggplot() +
  theme_classic(base_size = 20.0) +
  #ggtitle('P3 size:Local(low) - Local(high)' + theme(plot.title = element_text(hjust = 0.5))) +
  ylab(label = 'Scaled RT Slope') +
  scale_y_continuous(breaks = pretty_breaks()) +
  geom_bar(aes(y = scaled_rt_slope, x = learner2,fill =  as.factor(learner2)),data=df_corr_late,colour="black",fun.data = mean_sdl,stat = 'summary') +
  geom_beeswarm(aes(x = learner2,y = scaled_rt_slope),data=df_corr_late,dodge.width=0.9,cex=2.5) +
  geom_errorbar(aes(y = scaled_rt_slope, x = learner2),data=df_corr_late,size = 0.3,width = 0.2,fun.y = function(x) mean(x),fun.ymin = function(x) mean(x) - sd(x)/sqrt(length(x)),fun.ymax = function(x) mean(x) + sd(x)/sqrt(length(x)) ,stat = 'summary')+
  scale_x_discrete(name = 'Learner Group',labels = c("LDN","P3"))+   
  scale_fill_brewer(palette = 'Set2')+
  theme(legend.position = "none")

ggplot() +
  theme_classic(base_size = 20.0) +
  #ggtitle('P3 size:Local(low) - Local(high)' + theme(plot.title = element_text(hjust = 0.5))) +
  ylab(label = 'Accuracy') +
  scale_y_continuous(breaks = pretty_breaks()) +
  geom_bar(aes(y = accuracy, x = learner2,fill =  as.factor(learner2)),data=df_corr_late,colour="black",fun.data = mean_sdl,stat = 'summary') +
  geom_beeswarm(aes(x = learner2,y = accuracy),data=df_corr_late,dodge.width=0.9,cex=2.5) +
  geom_errorbar(aes(y = accuracy, x = learner2),data=df_corr_late,size = 0.3,width = 0.2,fun.y = function(x) mean(x),fun.ymin = function(x) mean(x) - sd(x)/sqrt(length(x)),fun.ymax = function(x) mean(x) + sd(x)/sqrt(length(x)) ,stat = 'summary')+
  scale_x_discrete(name = 'Learner Group',labels = c("LDN","P3"))+   
  scale_fill_brewer(palette = 'Set2')+
  theme(legend.position = "none")
```
