---
title: "ERP ERSP OC Word Analysis"
author: "JM Schneider"
date: "05/11/2022"
output: html_document
---

## Specialty Packages (first pass only)
```{r}
#install.packages("remotes")
#remotes::install_github("norment/normentR")
```

## Load Libraries
```{r}
library(stringr);
library(dplyr);
library(sjPlot);
library(sjmisc);
library(ggplot2);
library(interactions);
library(lme4);
library(afex);
library(tidyr);
library(normentR);
library(tableone);
```

## Load data
```{r}
ersp <- read.csv('all_ersp_data_new_ROIs.csv')
erp <- read.csv('erp_combined_new_ROIs.csv')
erp_subj <- read.csv('erp_subj_order.csv')

subj<-rep(erp_subj$subject_id, each = 40)
erp$subject_id<-rep(subj, 5)
#write.csv(erp,'erp_corrected_new_ROIs.csv')
```

## Keep only positive times in ERSP data & merge ERP/ERSP data
```{r}
ersp<-subset(ersp, timebin > 0)
#write.csv(ersp,'ersp_corrected_new_ROIs.csv')
```

## Read in Full DF
```{r}
df<-read.csv('ersp_erp_oc_df_new_ROIs.csv')
ast<-read.csv('WLNSF_jms_fieldtrip.csv')
df$subject_id<-str_remove(df$subject_id, "_WLNSF")
ast$PPVT.4.SS<-as.numeric(as.character(ast$PPVT.4.SS))
mat_ed<-read.csv('mat_ed_only.csv')
mat_ed$subject_id<-str_remove(mat_ed$subject_id, "_EXCLUDE")
ast$subject_id<-str_remove(ast$subject_id, "_EXCLUDE")
ast<-merge(ast, mat_ed, by = "subject_id")
```

# Splice time data into 200 msec time bins
```{r}
df$bins<-ifelse(df$timebin >=  100 & df$timebin <  300, "bin1",
            ifelse(df$timebin >=  300 & df$timebin <  500, "bin2", 
                 ifelse(df$timebin >=  500 & df$timebin <  700, "bin3", "bin10")))

df_bins<-df %>%
    group_by(bins,subject_id,cluster_ID) %>%
    summarise(across(everything(), mean))

df_bins %>%
  group_by(bins) %>%
  summarize(cor=cor(theta_pow, erp_combined))

df_bins <- merge(df_bins,ast, by = "subject_id")
df_bins$GORT_comp<-as.numeric(as.character((df_bins$GORT.V....comp)))
df_bins$C_ChronologicalAge<-scale(df_bins$ChronologicalAge, scale=FALSE)
df_bins$z_PPVT.4.SS<-scale(df_bins$PPVT.4.SS, scale=TRUE)
df_bins$z_GORT_comp<-scale(df_bins$GORT_comp, scale=TRUE)
df_bins$z_nwr<-scale(df_bins$NWR.PCC, scale=TRUE)
df_bins$z_twre<-scale(df_bins$TOWRE.TWRE, scale=TRUE)

df_bins$age_cat<-ifelse(df_bins$ChronologicalAge <  9, 8,
            ifelse(df_bins$ChronologicalAge >=  9 & df_bins$ChronologicalAge <  10,9, 
                 ifelse(df_bins$ChronologicalAge >= 10 & df_bins$ChronologicalAge <  11, 10,
                        ifelse(df_bins$ChronologicalAge >= 11 & df_bins$ChronologicalAge <  12, 11,
                              ifelse(df_bins$ChronologicalAge >= 12 & df_bins$ChronologicalAge <  13, 12,
                                    ifelse(df_bins$ChronologicalAge >= 13 & df_bins$ChronologicalAge <  14, 13,
                                          ifelse(df_bins$ChronologicalAge >= 14 & df_bins$ChronologicalAge <  15, 14,
                                                ifelse(df_bins$ChronologicalAge >= 15 & df_bins$ChronologicalAge <  16, 15,
                                                       ifelse(df_bins$ChronologicalAge >= 16 & df_bins$ChronologicalAge <  17, 16,"ageNA")))))))))

df_bins$age_cat<-as.factor(df_bins$age_cat)
df_bins$cluster_ID<-as.factor(df_bins$cluster_ID)

df_bins %>%
  group_by(cluster_ID, bins) %>%
  summarize(cor=cor(theta_pow, erp_combined))

bin1 <- subset(df_bins, bins == 'bin1')
bin2 <- subset(df_bins, bins == 'bin2')
bin3 <- subset(df_bins, bins == 'bin3')

mat_ed_count<-df_bins %>% group_by(subject_id) %>% tally(unique(mat_ed))
```

# RQ: For content words, what predicts the amplitude of the N400 and theta?
# Overview of results: The N400 and theta amplitude are not associated in the LC ROI. What appears to account for this is increasing age for theta, and increasing vocabulary knowledge for ERPs. 
#RQ sub-question1: N400 amplitude
```{r}
bin2$C_mat_ed<-scale(bin2$mat_ed, scale = FALSE)

lmer_erp_2<-lmer(erp_combined~ 1+C_mat_ed+theta_pow:cluster_ID+C_ChronologicalAge:cluster_ID+z_PPVT.4.SS:cluster_ID+z_GORT_comp:cluster_ID+z_nwr:cluster_ID+(1|subject_id), data=bin2)
summary(lmer_erp_2)

summary(lmer_erp_2)$coefficients  
#write.csv(summary(lmer_erp_2)$coefficients, 'erp_lmer_output.csv')
```
# Plot ERP Amplitude Relationships
```{r}
#svg("correlation_erp_ersp.svg")
plot_model(lmer_erp_2, type = "pred", terms = c("theta_pow","cluster_ID"))
#dev.off() 
```

#RQ sub-question2: Theta amplitude
```{r}
lmer_theta_2<-lmer(theta_pow~ 1+C_mat_ed+erp_combined:cluster_ID+C_ChronologicalAge:cluster_ID+z_PPVT.4.SS:cluster_ID+z_GORT_comp:cluster_ID+z_nwr:cluster_ID+ (1|subject_id), data=bin2)
summary(lmer_theta_2)

summary(lmer_theta_2)$coefficients  
#write.csv(summary(lmer_theta_2)$coefficients, 'ersp_lmer_output.csv')
```

## Exploratory (not included in manuscript)
## Plotting the time series data for ERPs across ages
```{r}
erp$subject_id<-str_remove(erp$subject_id, "_WLNSF")
erp<-merge(erp,ast, by = "subject_id" )

erp$age_cat<-ifelse(erp$ChronologicalAge <  10, "8-9",
            ifelse(erp$ChronologicalAge >=  10 & erp$ChronologicalAge <  12,"10-11", 
                 ifelse(erp$ChronologicalAge >= 12 & erp$ChronologicalAge <  14, "12-13",
                        ifelse(erp$ChronologicalAge >= 14 & erp$ChronologicalAge <  16, "14-15","ageNA"))))

erp_subset <- erp[c(1:5,20)]

erp_avg<-erp_subset %>%
    group_by(age_cat, timebin, cluster_ID) %>%
    summarise(across(everything(), mean))

erp_avg %>% 
  filter(cluster_ID %in% c("ROI_F"))  %>%
  ggplot( aes(x=timebin, y=erp_combined, group=age_cat, color=age_cat)) +
    geom_line()
```
## Plotting the time series data for ERPs  based on PPVT
```{r}
quantile(erp$PPVT.4.SS)
##   0%  25%  50%  75% 100% 
##  57   96  109  120  152 

erp$ppvt_cat<-ifelse(erp$PPVT.4.SS <=  96, "Low Vocabulary",
            ifelse(erp$PPVT.4.SS >=  120,"High Vocabulary", "Average Vocabulary"))

erp_subset <- erp[c(1:5,21)]

erp_avg<-erp_subset %>%
    group_by(ppvt_cat, timebin, cluster_ID) %>%
    summarise(across(everything(), mean))

erp_avg %>% 
  filter(cluster_ID %in% c("ROI_LC"))  %>%
  ggplot( aes(x=timebin, y=erp_combined, group=ppvt_cat, color=ppvt_cat)) +
    geom_smooth(method = "loess")

colors <- norment_pal(palette = "logo")(3)

ERPplot<-erp_avg %>% 
  filter(cluster_ID %in% c("ROI_LC"))  %>%
  ggplot( aes(x=timebin, y=erp_combined, group=ppvt_cat, color=ppvt_cat))  + 
  geom_line(size = 0.75) + 
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  scale_x_continuous(breaks = c(seq(-100,800,100))) + 
  #scale_y_continuous(breaks = c(seq(-.5,-,2))) + 
  coord_cartesian() +
  labs(x = "Time (ms)", 
       y = "Amplitude (µV)") + 
  theme_norment(ticks = TRUE, grid = FALSE) +
  theme(
    legend.position = c(0.9,0.1),
    axis.text.x = element_text(vjust = -0.1))
  
shift_axes(ERPplot, x = 0, y = 0)
#ERPplot

## Export data to plot in ERPlab
erp_subset <- erp[c(1,20,21)]
erp_export <-distinct(erp_subset,subject_id,.keep_all = TRUE)
write.csv(erp_export,"age_ppvt_group_ID.csv")
```

## Plotting the time series data for ERSPs
```{r}
ersp <- read.csv('all_ersp_data.csv')

ersp$subject_id<-str_remove(ersp$subject_id, "_WLNSF")
ersp<-merge(ersp,ast, by = "subject_id" )

ersp$age_cat<-ifelse(ersp$ChronologicalAge <  10, "8-9",
            ifelse(ersp$ChronologicalAge >=  10 & ersp$ChronologicalAge <  12,"10-11", 
                 ifelse(ersp$ChronologicalAge >= 12 & ersp$ChronologicalAge <  14, "12-13",
                        ifelse(ersp$ChronologicalAge >= 14 & ersp$ChronologicalAge <  16, "14-15","ageNA"))))

ersp_subset <- ersp[c(1:3,7,8,23)]

ersp_avg<-ersp_subset %>%
    group_by(age_cat, timebin, cluster_ID) %>%
    summarise(across(everything(), mean))

ersp_avg %>% 
  filter(cluster_ID %in% c("ROI_LC"))  %>%
  ggplot( aes(x=timebin, y=theta_pow, group=age_cat, color=age_cat)) +
    geom_line()
```

## Create a Table of Demographic Data sorted by Age Groups
```{r}
ast<-merge(ast,mat_ed_count, by = "subject_id")
ast$age_cat<-ifelse(ast$ChronologicalAge <  10, "8-9",
            ifelse(ast$ChronologicalAge >=  10 & ast$ChronologicalAge <  12,"10-11", 
                 ifelse(ast$ChronologicalAge >= 12 & ast$ChronologicalAge <  14, "12-13",
                        ifelse(ast$ChronologicalAge >= 14 & ast$ChronologicalAge <  16, "14-15","ageNA"))))

ast$GORT.V....comp<-as.numeric(as.character(ast$GORT.V....comp))
ast$mat_ed<-as.factor(ast$mat_ed)

myvars <- c("ChronologicalAge","PPVT.4.SS","GORT.V....comp","NWR.PCC", "mat_ed")

table2 <- CreateTableOne(myvars, ast, strata = c("age_cat"))
table2
#write.table(table2, file = "demo_table.txt", sep = ",", quote = FALSE, row.names = T)

```
