---
title: "Regression Learning Manuscript"
author: "Julie M. Schneider"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/', echo=TRUE, tidy=TRUE, message=F, warning=F, cache=T)
```

## Library Import
```{r,include=FALSE,echo=FALSE,warning=FALSE}
#Loading library
library(ez); library(ggplot2); library(multcomp); library(nlme); library(pastecs); library(reshape)
library(Hmisc); library (WRS2); library(lme4); library(lmerTest); library(plyr); library(splitstackshape) 
library(naniar); library(tidyr); library(dplyr); library(memisc); library(psychReport); library(foreign); library(interactions); library(tableone); library(officer); library(magrittr); library(corrplot); library(effectsize); library(Rcmdr);library(scales)

julie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom graphics theme for ggplot2
```

## Import and merge data
```{r,include=FALSE,echo=FALSE,warning=FALSE}
setwd("/Volumes/GoogleDrive/My Drive/Projects/FLUX20/")
df= read.csv("Multiple_Regression_120920_data.csv")
```

### Create column for Age
```{r}
df$age_year<- ifelse(df$age_months < 108, 8, ifelse(df$age_months >= 108 & df$age_months < 120, 9, ifelse(df$age_months >= 120 & df$age_months < 132, 10, ifelse(df$age_months >= 132 & df$age_months < 144, 11, ifelse(df$age_months >= 144 & df$age_months < 156, 12, ifelse(df$age_months >= 156 & df$age_months < 168,13, ifelse(df$age_months >= 168 & df$age_months < 180, 14,ifelse(df$age_months >= 180 & df$age_months < 192, 15,ifelse(df$age_months > 192, 16,99)))))))))
```

### Use most complete dataset
```{r}
df$subject<-as.factor(as.character(df$subject))
df<- df[-c(2,45,58,62,63,73,76,85,86,88,105,140,207,223,237,238,256,261,262,266,268,275),]
```

## Evaluate the existing dataset & remove outliers
```{r}
df$GORT_ORI<-as.numeric(as.character(df$GORT_ORI))
df$rate<-as.numeric(as.character(df$rate))
df$acc<-as.numeric(as.character(df$acc))
df$fluency<-as.numeric(as.character(df$fluency))
df$comp<-as.numeric(as.character(df$comp))
df$PPVT_SS<-as.numeric(as.character(df$PPVT_SS))
df$Mat_ed<-as.numeric(as.character(df$Mat_ed))
df$NWR.PCC<-as.numeric(as.character(df$NWR.PCC))
df$twre_pde<-as.numeric(as.character(df$twre_pde))
df$digit<-as.numeric(as.character(df$digit))
df$lang_exp<-as.numeric(as.character(df$lang_exp))
df$subject<-as.numeric(as.character(df$subject))
df$WL_PC<-as.numeric(as.character(df$WL_PC))

range(df$WL_PC)
df <- df[!(df$WL_PC < .08), ]

#df<- mutate_all(df, function(x) as.numeric(as.character(x)))
#colMeans(df)
```

## RQ: Do SES related gaps in word learning change across ages 8-15?
### Answer: No, there are significant SES related gaps in WL (model 2) but these do not significantly change on the basis of age (model 3; cross-sectional data)
```{r}
onePredictorModel <- lm(WL_PC ~ age_months + lang_exp, df)
twoPredictorModel <- lm(WL_PC ~ age_months + lang_exp + Mat_ed, df)
threePredictorModel <- lm(WL_PC ~ lang_exp +age_months*Mat_ed, df)

summary(onePredictorModel)
effectsize(onePredictorModel)
summary(twoPredictorModel)
effectsize(twoPredictorModel)
summary(threePredictorModel)
effectsize(threePredictorModel)

anova(onePredictorModel,twoPredictorModel,threePredictorModel)

#pdf("Figs/mated_regression.pdf")
interact_plot(threePredictorModel, pred = age_months, modx = Mat_ed, plot.points = TRUE)
#dev.off()
```
## RQ: What variables are strongly associated with word learning ability?
### Correlation between Languag and Environmental Measures
```{r}
myvars <- c("WL_PC","PPVT_SS","Mat_ed","age_months","lang_exp","digit","comp", "NWR.PCC","twre_pde")
df_subset <- df[myvars]

#Overall Correlation
overall_corr<-rcorr(as.matrix(df_subset))
corrplot(overall_corr$r, method = "color", type = "upper",addCoef.col="dark grey", tl.col = "black", p.mat = overall_corr$P, sig.level = 0.05, insig = "blank", tl.cex = 0.9)

#Partial Correlation, where each bivariate comparison controls for all other variables in matrix
par_overall_corr <- partial.cor(df_subset, tests=TRUE, use=c("pairwise.complete.obs"))
#pdf(file = "partial_corr_behavior_adjusted.pdf")
corrplot(par_overall_corr$R, method = "color", type = "upper",addCoef.col="black", tl.col = "black", p.mat = par_overall_corr$P.unadj, sig.level = 0.05, insig = "blank", tl.cex = 0.9)

#dev.off()

#Correlation between specific variables, while controlling for a specific variable
source("pcor.R")
pcor.test(df$PPVT_SS, df$age_months, df$Mat_ed, use = c("mat"), method = c("pearson"), na.rm = T)
pcor.test(df$PPVT_SS, y = df$age_months, 
          z = df[, c("Mat_ed", "lang_exp", "comp")])
```

## RQ: Which abilities are most strongly related with an individual's word learning? SES, age, vocabulary, language exposure, reading comprehension, working memory, and phonological working memory 
### Answer: Age, vocabulary, and reading comprehension are all strongly positively associated. Phonological working memory is 0.06 (will not investigate further).
```{r}
myvars <- c("WL_PC","PPVT_SS","Mat_ed","age_months","lang_exp","digit","comp", "NWR.PCC")
df_step <- df[myvars]
df_step<-na.omit(df_step)

library(MASS)
library(car)
library(lm.beta)
# Fit the full model 
full.model <-lm(WL_PC ~ lang_exp + digit + NWR.PCC + comp+ Mat_ed + PPVT_SS + age_months, df_step)

summary(full.model)
lm.beta(full.model)
effectsize(full.model)

# Stepwise regression model
# step.model <- stepAIC(full.model, direction = "backward", 
#                       trace = FALSE)
# 
# summary(step.model)
# lm.beta(step.model)
# effectsize(step.model)
```

### Plot Vocabulary regression
```{r}
#jpeg("Figs/vocab_regression_022321.jpeg")
ggplot(df, aes(x=PPVT_SS, y = WL_PC)) + 
  labs(title="Vocabulary & Word Learning Ability", y = "Word Learning Accuracy", x = "Vocabulary") + 
  geom_point(size=2,color='#E69F00') + 
  julie.theme+ 
  theme(plot.title = element_text(hjust = 0.5, size =20), axis.text = element_text(size = 20), axis.title = element_text(size=20)) +
  geom_smooth(method=lm,  linetype="dashed",
             color="black")
#dev.off()
```
### Plot comp regression
```{r}
#jpeg("Figs/vocab_regression.jpeg")
ggplot(df, aes(x=comp, y = WL_PC)) + labs(title="Reading Comprehension predicts Word Learning Ability", y = "Word Learning Accuracy", x = "Reading Comprehension") + geom_point(size=2,color='#E69F00') + julie.theme+ theme(plot.title = element_text(hjust = 0.5)) + geom_smooth(method=lm,  linetype="dashed",
             color="black")
#dev.off()
```

## RQ: Of the variables known to be important to word learning, do they vary across SES or age -- things known to impact word learning ability?
## Vocabulary: No.

```{r}
vocab_model <-lm(WL_PC ~ lang_exp + digit + NWR.PCC + comp+  Mat_ed * PPVT_SS * age_months, df_step)

summary(vocab_model)
lm.beta(vocab_model)
effectsize(vocab_model)

step.model <- stepAIC(vocab_model, direction = "backward",
                      trace = FALSE)
```

## Reading Comp: YES. Simple slope analyses indicate that as children get older they perform significantly better on the word learning task. However, for middle and lower SES children under the age of 12, better reading comprehension abilities are positively associated with better word learning outcomes. 
```{r}
comp_model <-lm(WL_PC ~ lang_exp + digit + NWR.PCC + PPVT_SS +  Mat_ed * comp * age_months, df)

summary(comp_model)
lm.beta(comp_model)
effectsize(comp_model)


young <- df[(df$age_year <= 9),]
midage <- df[(df$age_year > 9 & df$age_year <= 12),]
old <- df[(df$age_year >= 13),]
lowses <- df[(df$ses_group == "Low SES"),]
highses <- df[(df$ses_group == "High SES"),]

young_model <-lm(WL_PC ~   Mat_ed * comp, midage)
summary(young_model)
lm.beta(young_model)
effectsize(young_model)

sim_slopes(comp_model, pred = comp, modx = age_months, mod2 = Mat_ed , johnson_neyman = TRUE)

```

### Plot reading interaction
```{r}
interact_plot(comp_model, pred = comp, modx = Mat_ed, mod2 = age_months, plot.points = TRUE)
```

```{r}
df$ses_age_group<- ifelse(df$age_year <= 9 & df$Mat_ed <= 1, "Young Low SES", 
                          ifelse(df$age_year > 9 & df$age_year <= 12 & df$Mat_ed <= 1, "Mid-Age Low SES", 
                                 ifelse(df$age_year >= 13 & df$Mat_ed <= 1, "Older Low SES", 
                                        ifelse(df$age_year <= 9 & df$Mat_ed == 2, "Young Mid SES", 
                                               ifelse(df$age_year > 9 & df$age_year <= 12 & df$Mat_ed == 2, "Mid-Age Mid SES", 
                                                      ifelse(df$age_year >= 13 & df$Mat_ed == 2, "Older Mid SES", 
                                                             ifelse(df$age_year <= 9 & df$Mat_ed >= 3, "Young High SES",
                                                                  ifelse(df$age_year > 9 & df$age_year <= 12 & df$Mat_ed >= 3, "Mid-Age High SES",
                                                                         ifelse(df$age_year >= 13 & df$Mat_ed >= 3, "Older High SES",99)))))))))
```

```{r}
df$ses_age_group_2<- ifelse(df$age_year <= 9 & df$Mat_ed <= 2, "Young Low SES", 
                          ifelse(df$age_year > 9 & df$age_year <= 12 & df$Mat_ed <= 2, "Mid-Age Low SES", 
                                 ifelse(df$age_year >= 13 & df$Mat_ed <= 2, "Older Low SES", 
                                       ifelse(df$age_year <= 9 & df$Mat_ed >= 3, "Young High SES",
                                            ifelse(df$age_year > 9 & df$age_year <= 12 & df$Mat_ed >= 3, "Mid-Age High SES",
                                                      ifelse(df$age_year >= 13 & df$Mat_ed >= 3, "Older High SES",99))))))
```

```{r}
df$ses_group_2 <- ifelse(df$Mat_ed <=2, "Low SES", "High SES")
```

## Group Means
```{r}
df %>%
  group_by(ses_age_group_2) %>%
  summarise_at(vars(WL_PC,PPVT_SS,comp), funs(mean(., na.rm=TRUE)))
```

## New Plotting Approach (not based on models)
```{r}
# svg(filename="comp_ellipse.svg", 
#     width=7, 
#     height=5 , 
#     pointsize=12)

p <- ggplot(df, aes(comp,WL_PC, color = ses_age_group_2))+
  geom_point() + 
  scale_color_manual(values = c("royalblue", "brown2","darkblue","maroon","lightblue", "lightsalmon")) +
  stat_ellipse()
p + stat_ellipse()

#dev.off()
```

```{r}
# svg(filename="vocab_ellipse.svg", 
#     width=7, 
#     height=5 , 
#     pointsize=12)

scatterPlot <- ggplot(df,aes(PPVT_SS, WL_PC, color=ses_age_group_2)) + 
  geom_point() + 
  scale_color_manual(values = c("royalblue", "brown2","darkblue","maroon","lightblue", "lightsalmon")) +
  stat_ellipse()
scatterPlot

# dev.off()
```

## New Plotting Apprach (based on models)
```{r}
jpeg("Figs/reading_regression_022321.jpeg")
predicted_df <- data.frame(WL_PC_pred = predict(comp_model, df), comp=df$comp, ses_age_group_2=df$ses_age_group_2)

ggplot(predicted_df, aes(comp,WL_PC_pred, color = ses_age_group_2))+
  geom_point() + 
  julie.theme +
 scale_color_manual("SES x Age Group",values = c("royalblue", "brown2","darkblue","maroon","lightblue", "lightsalmon")) +
   xlab("Reading Comprehension") + ylab("Word Learning Accuracy") +stat_smooth(method=lm, se =FALSE) +theme(text = element_text(size = 15))   
dev.off()
```

## New Plotting Apprach (based on models)
```{r}
predicted_df <- data.frame(WL_PC_vocab_pred = predict(vocab_model, df), vocab=df$PPVT_SS, ses_age_group_2=df$ses_age_group_2)

ggplot(predicted_df, aes(vocab,WL_PC_vocab_pred, color = ses_age_group_2))+
  geom_point() + 
  julie.theme +
  scale_color_manual("SES x Age Group",values = c("royalblue", "brown2","darkblue","maroon","lightblue", "lightsalmon")) +
  xlab("Vocabulary") + ylab("Word Learning Accuracy")

#stat_ellipse() +
```

## New Plotting Apprach (based on models)
```{r}
predicted_df <- data.frame(WL_PC_vocab_pred = predict(vocab_model, df), vocab=df$PPVT_SS, ses_age_group_2=df$ses_group_2)

ggplot(predicted_df, aes(vocab,WL_PC_vocab_pred, color = ses_age_group_2))+
  geom_point() + 
  scale_color_manual("SES Group",values = c("royalblue", "brown2")) +
  xlab("Vocabulary") + ylab("Word Learning Accuracy") 
```

## Create a Table of Demographic Data sorted by Maternal Education
```{r}
table2 <- CreateTableOne(myvars, df, strata = c("Mat_ed"))
table2
#write.table(table2, file = "demo_stats.txt", sep = ",", quote = FALSE, row.names = T)

mean(df$lang_exp)
length(which(df$lang_exp == 1))
mean(df$Ethnicity)
length(which(df$Ethnicity ==0))
t.test(df$lang_exp~df$ses_group)

#Number of bilingual and monolingual kids per Mat Ed group
df %>% group_by(Mat_ed,lang_exp) %>% tally()
#Race Breakdown
df %>% group_by(Race) %>% tally()
```





